\documentclass[12pt]{article}

\usepackage{geometry}   % To control page margins
\usepackage{fancyhdr}   % To customize the header and footer
\usepackage{enumitem}   % For better lists
\usepackage{setspace}   % For line spacing
\usepackage{parskip}    % For better paragraph spacing

\geometry{a4paper, margin=1in}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Operating Systems} 
\fancyfoot[C]{\thepage}

\title{Operating System Services}
\author{0xSaad / Saad Almalki}
\date{\today}

\begin{document}

\maketitle

\section{OS Services}
There's many services provided by the OS to the users and to the programs.
\section{User Services}
\begin{itemize}


    \item User interface : Most Operating Systems have a user interface, CLI or GUI or even both and touchscreen UI.
    \item Program execution : load program into memory and run it and terminate the processes.
    \item I/O operations : May program require I/O Device to perform operations.
    \item File-system manipulation : The system must be able to make files, directories, read, write, and delete them.
    \item Communication : Processes may exchange information , In the same computer or between computers over a network.
    \item Accounting : Keep track of resource usage by users.
    \item Error detection : Continues monitor the systrm to detect potential errors.
    \item Process and security : Allows and deny users , Manages the control access to the system.
    \item Resource allocating : Allocate resources to processes as needed.
\end{itemize}

\section{System Calls}
Programming interface to services provided by the OS ,And can be accessed via High-level API (Application Programming Interface).

Most details of OS interface hidden from programmer by API.

\subsection{Types of System Calls}
\begin{itemize}
    \item Process Control : end, abort, load, execute.
    \item File Management : create, delete, open, close, read, write.
    \item Device Management : request device, release device, read, write.
    \item Information Maintenance : get time or date, get system data, set time or date.
    \item Communication : create message, send message, receive message.
    \item Protection : control access to resources.
\end{itemize}

\subsection{System Calls Examples}
\begin{itemize}
    \item Process Control : \textbf{fork(), exec(), wait(), exit()}
    \item File Management : \textbf{open(), read(), write(), close()}
    \item Device Management : \textbf{read(), write()}
    \item Information Maintenance : \textbf{sleep(), alarm()}
    \item Communication : \textbf{pipe()}
    \item Protection : \textbf{chmod() , chown() , umask()}
\end{itemize}

\section{System Programs}
System programs provides convenient environment for development and execute.

Types of System Programs :
\begin{itemize}
    \item File management : copy, delete, rename, print, dump, list files.
    \item Status information : date, time, system status.
    \item File modification : text editors.
    \item Programming language support : compilers, assemblers, debuggers.
    \item Program loading and execution : load and run programs.
    \item Communications : send, receive messages, use internet protocols.
    \item Background services : print spooling, file servers, network management.
    \item Application programs : word processors, spreadsheets, databases, video games. (Not part of OS and run by users)
\end{itemize}

\begin{flushright}
Created by: Saad Almalki \\
\end{flushright}

\end{document}
